Manuscript 1 code

### 2024-10-02 manuscript_1_code.Rmd
completed figure 3 panel

#don't need to use this, was just to check both methods
#fg_summary_site_AVG <- macro_functional_groups %>%
  #group_by(site, year, habitat, functional_group) %>%
  #summarise(mean_fg_cover = mean(prop_cover))

#fg_summary %>%
  #filter(sum_fg_cover==1)

@@

fg_summary_site_wide <- fg_summary_site_wide_SUM %>%
   mutate(site_habitat = paste(site, habitat, sep = "_"))
```

@@

```{r figure 3b plot}
(figure_3b <- 
  ggplot() +
  geom_point(data = functional_richness_quad_cover,
             aes(x = fg_richness_quad, y = cover_trans, colour = habitat), size = 1, alpha = 0.005) + 
  geom_line(data = filtered_cover_effects, aes(x = fg_richness_quad, y = inv_logit(yvar), colour = habitat)) +
  geom_ribbon(data = filtered_cover_effects, aes(x = fg_richness_quad, ymin = inv_logit(LCL), ymax = inv_logit(UCL), fill = habitat),
              alpha = 0.3) +
  scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
  labs(y = "", x = "Functional Richness", title = "3b.") +
  theme_bw() + 
  theme(axis.title.x = element_text(size=24), 
        axis.title.y = element_text(size=24), 
        axis.text.x = element_text(size=22), 
        axis.text.y = element_text(size=22),
        legend.text = element_text(size=22),
        plot.title = element_text(size=26), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_rect(fill = "white"),
        legend.title = element_blank(),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black"))  # add box around legend
         )

@@

@@

>>>>

@@

fg_summary_site_wide <- fg_summary_site_wide_SUM 
```

@@

```{r figure 3b plot}
(figure_3b <- 
  ggplot() +
  geom_point(data = functional_richness_quad_cover,
             aes(x = fg_richness_quad, y = cover_trans, colour = habitat), size = 1, alpha = 0.005) + 
  geom_line(data = filtered_cover_effects, aes(x = fg_richness_quad, y = inv_logit(yvar), colour = habitat)) +
  geom_ribbon(data = filtered_cover_effects, aes(x = fg_richness_quad, ymin = inv_logit(LCL), ymax = inv_logit(UCL), fill = habitat),
              alpha = 0.3) +
  scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
  labs(y = "", x = "", title = "3b.") +
  theme_bw() + 
  theme(axis.title.x = element_text(size=24), 
        axis.title.y = element_text(size=24), 
        axis.text.x = element_text(size=22), 
        axis.text.y = element_text(size=22),
        legend.text = element_text(size=22),
        plot.title = element_text(size=26), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_rect(fill = "white"),
        legend.title = element_blank(),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black"))  # add box around legend
         )

@@

## Figure 3e - functional richness at site level
### model

```{r site level fg richness model}
#merge site level cover with site functional group richness

functional_richness_site_cover<- merge(alpha_diversity_site_macro_with_cover, fg_summary_site_wide,
                                by = c("year", "site",  "habitat"))

site_cover_mod_fg <- glmmTMB(cover_trans ~ fg_richness_site*habitat + (1|site_habitat) + (1|year), family = beta_family(), data = functional_richness_site_cover)

summary(site_cover_mod_fg)
Anova(site_cover_mod_fg)

#                            Chisq Df Pr(>Chisq)    
# fg_richness_site         30.9412  1   2.66e-08 ***
# habitat                   4.8929  3     0.1798    
# fg_richness_site:habitat  4.2193  3     0.2387

hist(residuals(site_cover_mod_fg)) # positive skew
plot(residuals(site_cover_mod_fg) ~ fitted(site_cover_mod_fg)) # a little heteroscedastic but can't be less than 0
r.squaredGLMM(site_cover_mod_fg) # 0.180, 0.563
emtrends(site_cover_mod_fg, pairwise ~ habitat, var = "fg_richness_site") # no difference between the habitats

```
```{r site fg richness cover plot setup}
diversity_ranges_site <- functional_richness_site_cover %>%
  group_by(habitat) %>%
  summarise(min_fg_richness_site = min(fg_richness_site),
            max_fg_richness_site = max(fg_richness_site),
            min_cover= min(cover_trans),
            max_cover = max(cover_trans))

site_cover_emm <- emmip(site_cover_mod_fg,
                        habitat ~ fg_richness_site,
                        at = list(fg_richness_site = seq(0,17,1)), plotit = F, CIs = T)

filtered_site_cover_effects <- site_cover_emm %>%
  left_join(diversity_ranges_site, by = c("habitat")) %>%
  filter(fg_richness_site >= min_fg_richness_site & fg_richness_site <= max_fg_richness_site)
```

### 3e plot

```{r figure 3e plot }
(figure_3e <- 
  ggplot() +
  geom_point(data = functional_richness_site_cover,
             aes(x = fg_richness_site, y = cover_trans, colour = habitat), size = 1, alpha = 0.3) + 
  geom_line(data = filtered_site_cover_effects, aes(x = fg_richness_site, y = inv_logit(yvar), colour = habitat),
            size = 1.5) +
  geom_ribbon(data = filtered_site_cover_effects, aes(x = fg_richness_site, ymin = inv_logit(LCL), ymax = inv_logit(UCL), fill = habitat),
              alpha = 0.3) +
  scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
  labs(y = "", x = "Functional Richness", title = "3e.") +
  theme_bw() + 
  theme(axis.title.x = element_text(size=24), 
        axis.title.y = element_text(size=24), 
        axis.text.x = element_text(size=22), 
        axis.text.y = element_text(size=22),
        legend.text = element_text(size=22),
        plot.title = element_text(size=26), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_rect(fill = "white"),
        legend.title = element_blank(),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black"))  # add box around legend
         )
#ggsave(filename = "output/figure_3e_v2.png", figure_3e, height = 10, width = 14)
```

@@

## Figure 3 Full Plot
```{r ggpubr fig3, fig.height = 8, fig.width=16}

(figure_3 <- ggarrange(figure_3a, figure_3b, figure_3c, figure_3d, figure_3e, figure_3f, 
                      ncol = 3, nrow = 2, 
                      common.legend = TRUE,
                      legend = "bottom", 
                      font.label = list(size = 20, color = "black", face = "plain")))

# ggsave(filename = "output/figure_3.png", figure_3, height = 8, width = 16)
```
